---
title: "Computing Assignment 2"
author: "Angela Huynh"
date: "September 20, 2016"
output: pdf_document
---

**Grade: 53/50**


1. Working with data: In the datasets folder on the course GitHub repo, you will find a file called cancer.csv, which is a dataset in comma-separated values (csv) format. This is a large cancer incidence dataset that summarizes the incidence of different cancers for various subgroups.

i) Load the data set into R and make it a data frame called cancer.df

```{r}
#cancer.df <- read.csv("~/Documents/Vanderbilt 1/Semester 1/BIOS 6301/Assignments/cancer.csv")
cancer.df <- read.csv("cancer.csv")
#View(cancer.df)
```
ii) Determine the number of rows and columns in the data frame.

```{r}
nrow(cancer.df) #number of rows
ncol(cancer.df) #number of columns
```

iii) Extract the names of the columns in cancer.df.

```{r}
colnames(cancer.df)
```

iv) Report the value of the 3000th row in column 6. 

```{r}
cancer.df[3000,6]
```

v) Report the contents of the 172nd row.

```{r}
cancer.df[172,]
```

vi) Create a new column that is the incidence rate (per 100,000) for each row.

```{r}
cancer.df$incidencerate <- cancer.df$incidence/100000
```

**JC Grading - 1**
For incidence rate above should be incidence / population * 100000


vii) How many subgroups (rows) have a zero incidence rate? 

```{r}
length(which(cancer.df$incidencerate==0))
```

viii) Find the subgroup with the highest incidence rate.

```{r}
which.max(cancer.df$incidencerate)
```


**JC Grading - 1**
syntax is fine but answer is incorrect b/c of how incidence rate was calculated


2. Data types 

i) Create the following vector: x <- c("5","12","7"). Which of the following commands will produce an error message? For each command, Either explain why they should be errors, or explain the non-erroneous result. 

```{r}
x <- c("5","12","7")
max(x) #max is listed as "7" 
sort(x) #sorts as "12" "5" "7" because the function looks at the first character in the string
#sum(x) #produces an error because these are not numeric types, thus they cannot be summed
```

ii) For the next two commands, either explain their results, or why they should produce errors.

```{r}
y <- c("5",7,12) 
y #output is the same as x
#y[2] + y[3] #cannot add strings together
```

iii) For the next two commands, either explain their results, or why they should produce errors.

```{r}
z <- data.frame(z1="5",z2=7,z3=12)
z
z[1,2] + z[1,3] #adds 7 and 12 together, getting 19. Integers are able to be added. 
```

3. Data structures: Give R expressions that return the following matrices and vectors (i.e. do not construct them manually).

i) $(1,2,3,4,5,6,7,8,7,6,5,4,3,2,1)$

```{r}
seq1 <- seq(from=1, to=8, by=1) 
seq2 <- seq(from=7, to=1, by=-1)
c(seq1,seq2)
```

ii) $(1,2,2,3,3,3,4,4,4,4,5,5,5,5,5)$

```{r}
c <- vector()
for (i in (1:5)){
  c <- append(c, rep(i,i))
}
c
```

iii) $\begin{pmatrix} 0 & 1 & 1 \ 1 & 0 & 1 \ 1 & 1 & 0 \ \end{pmatrix}$

```{r}
a <- matrix(rep(1,9),3,3)
`diag<-`(a,0)
```

iv) $\begin{pmatrix} 1 & 2 & 3 & 4 \\ 1 & 4 & 9 & 16 \\ 1 & 8 & 27 & 64  \\ 1 & 16 & 81 & 256 \\ 1 & 32 & 243 & 1024  \\ \end{pmatrix}$

```{r}

(b <- matrix(c(seq(1,4),seq(1,4)^2,seq(1,4)^3,seq(1,4)^4,seq(1,4)^5), nrow=5, ncol=4, byrow=TRUE))

```

4.Basic Programming

i) Let $h(x,n)=1+x+x^2+\ldots+x^n = \sum_{i=0}^n x^i$. Write an R program to calculate $h(x,n)$ using a for loop.

```{r}
hx <- function(x,n){
  valuei <- 0
  sumx <- 0
  for (i in(0:n)){
    sumx <- sumx + x^(i)
    valuei <- valuei + 1
  }
  return(sumx)
}
```

ii) If we list all the natural numbers below 10 that are multiples of 3 or 5, we get 3, 5, 6 and 9. The sum of these multiples is 23. Write an R program to perform the following calculations.

  a) Find the sum of all the multiples of 3 or 5 below 1,000.
  
```{r}
multiple35 <- 0
count <- 0
for (i in(1:999)){
  if (i%%3==0 | i%%5==0){
    count <- count + i
    multiple35 <- multiple35 + i
  }
}
multiple35
```
  
  b) Find the sum of all the multiples of 4 or 7 below 1,000,000.
  
```{r}
multiple47 <- 0
count1 <- 0
for (i in(1:999999)){
  if (i%%4==0 | i%%7==0){
    count1 <- count1 + i
    multiple47 <- multiple47 + i
  }
}
multiple47
```

iii) Each new term in the Fibonacci sequence is generated by adding the previous two terms. By starting with 1 and 2, the first 10 terms will be $(1, 2, 3, 5, 8, 13, 21, 34, 55, 89)$. Write an R program to calculate the sum of the first 15 even-valued terms.

```{r}
countevens <- 1
sumevenfibs <- 2
fibterms <- vector()
evenfib <- vector()
fibterms[1] <- 1
fibterms[2] <- 2
evenfib[1] <- 2

for (i in(3:100)){
  fibterms[i] <- fibterms[i-1] + fibterms[i-2]
  if (fibterms[i]%%2==0){
    evenfib <- append(evenfib, fibterms[i])
    sumevenfibs <- sumevenfibs + fibterms[i]
    countevens <- countevens + 1
    if (countevens >= 15){
      break
    }
  }
}

sumevenfibs
```

**JC Grading +5 Bonus**
