---
title: "Homework 6"
author: "Angela Huynh"
date: "12/01/2016"
output: pdf_document
---
```{r, include=FALSE, cache=TRUE, tidy=TRUE}
install.packages("ggplot2", repos = "http://cran.us.r-project.org")
library(ggplot2)
install.packages("testthat", repos = "http://cran.us.r-project.org")
library(testthat)
```

# Question 1

Consider the following very simple genetic model (very simple -- don't worry if you're not a geneticist!). A population consists of equal numbers of two sexes: male and female. At each generation men and women are paired at random, and each pair produces exactly two offspring, one male and one female. We are interested in the distribution of height from one generation to the next. Suppose that the height of both children is just the average of the height of their parents, how will the distribution of height change across generations?

Represent the heights of the current generation as a dataframe with two variables, m and f, for the two sexes. We can use rnorm to randomly generate the population at generation 1:

```{r, cache=TRUE, tidy=TRUE}
# randomly generate population heights at generation 1
pop1 <- data.frame(m = rnorm(100, 160, 20), f = rnorm(100, 160, 20))
```

The following function takes the data frame pop and randomly permutes the ordering of the men. Men and women are then paired according to rows, and heights for the next generation are calculated by taking the mean of each row. The function returns a data frame with the same structure, giving the heights of the next generation.

```{r, cache=TRUE, tidy=TRUE}
# returns heights of the next generation
next_gen <- function(pop) {
    pop$m <- sample(pop$m)
    pop$m <- rowMeans(pop)
    pop$f <- pop$m
    pop
}
```

Use the function next_gen to generate nine generations (you already have the first), then use the function hist to plot the distribution of male heights in each generation (this will require multiple calls to hist). The phenomenon you see is called regression to the mean. Provide (at least) minimal decorations such as title and x-axis labels.

```{r, cache=TRUE, tidy=TRUE}
# create generations 1-9
# 1: pop
pop2 <- next_gen(pop1)
pop3 <- next_gen(pop2)
pop4 <- next_gen(pop3)
pop5 <- next_gen(pop4)
pop6 <- next_gen(pop5)
pop7 <- next_gen(pop6)
pop8 <- next_gen(pop7)
pop9 <- next_gen(pop8)

# plot distribution of male heights in each generation
par(mfrow=c(3,3))
hist(x = pop1$m, xlab = "Height", main = "Histogram of Gen 1 Males", col = "aquamarine")
hist(x = pop2$m, xlab = "Height", main = "Histogram of Gen 2 Males", col = "coral 1")
hist(x = pop3$m, xlab = "Height", main = "Histogram of Gen 3 Males", col = "cornflower blue")
hist(x = pop4$m, xlab = "Height", main = "Histogram of Gen 4 Males", col = "lavender")
hist(x = pop5$m, xlab = "Height", main = "Histogram of Gen 5 Males", col = "maroon")
hist(x = pop6$m, xlab = "Height", main = "Histogram of Gen 6 Males", col = "dark olive green")
hist(x = pop7$m, xlab = "Height", main = "Histogram of Gen 7 Males", col = "light sky blue")
hist(x = pop8$m, xlab = "Height", main = "Histogram of Gen 8 Males", col = "orange red 2")
hist(x = pop9$m, xlab = "Height", main = "Histogram of Gen 9 Males", col = "violet")
```

```{r, include=FALSE}
par(mfrow=c(1,1))
```

# Question 2

Use the simulated results from question 1 to reproduce (as closely as possible) the following plot in ggplot2.

```{r, cache=TRUE, tidy=TRUE}
idx <- sort(rep(1:9, nrow(pop1))) # create index for respective generation group
gen.data <- cbind(rbind(pop1, pop2, pop3, pop4, pop5, pop6, pop7, pop8, pop9), idx)
ggplot(data = gen.data) + geom_point(mapping=aes(x=m, y=f, alpha=0.005, shape='16')) + facet_wrap( ~idx)
```

# Question 3

You calculated the power of a study design in question #2 of assignment 3. The study has two variables, treatment group and outcome. There are two treatment groups (0, 1) and they should be assigned randomly with equal probability. The outcome should be a random normal variable with a mean of 60 and standard deviation of 20. If a patient is in the treatment group, add 5 to the outcome.

Starting with a sample size of 250, create a 95% bootstrap percentile interval for the mean of each group. Then create a new bootstrap interval by increasing the sample size by 250 until the sample is 2500. Thus you will create a total of 10 bootstrap intervals. Each bootstrap should create 1000 bootstrap samples. (4 points)

```{r, cache=TRUE, tidy=TRUE}
set.seed(100) # set seed to replicate outcome
n <- 1000 # number of times we repeat procedure
alpha <- .05

# function to get treatment and outcome data
get.data <- function(sample.size) {
  for(i in (1:sample.size)) {
  treatment <- rbinom(n=sample.size, size=1, prob=.5)
  outcome <- rnorm(n=sample.size, mean=60, sd=20)
  for(a in (1:sample.size)) {
    if (treatment[a] == 1) {
      outcome[a] <- outcome[a] + 5
      }
    }
  }
  study.data <- data.frame(cbind(treatment, outcome))
  return(study.data)
}

# vector of means for two treatment groups

# treatment = 0 
means.group0 <- vector()
lower.bounds0 <- vector()
upper.bounds0 <- vector()

# treatment = 1
means.group1 <- vector()
lower.bounds1 <- vector()
upper.bounds1 <- vector()

# bootstrapping 
for (i in seq(250,2500,250)) {
  current.data <- get.data(i)
  means.0 <- vector()
  means.1 <- vector()
  for (j in 1:n) {
    sample.data <- current.data[sample(nrow(current.data), size = i, replace = TRUE), ]
    means.0[j] <- mean(sample.data[which(sample.data$treatment==0),]$outcome)
    means.1[j] <- mean(sample.data[which(sample.data$treatment==1),]$outcome)
  }
  means.group0 <- c(means.group0, mean(means.0))
  means.group1 <- c(means.group1, mean(means.1))
  lower0 <- quantile(means.0, c(0.025, 0.975))[1]
  upper0 <- quantile(means.0, c(0.025, 0.975))[2]
  lower1 <- quantile(means.1, c(0.025, 0.975))[1]
  upper1 <- quantile(means.1, c(0.025, 0.975))[2]
  lower.bounds0 <- c(lower.bounds0, lower0)
  upper.bounds0 <- c(upper.bounds0, upper0)
  lower.bounds1 <- c(lower.bounds1, lower1)
  upper.bounds1 <- c(upper.bounds1, upper1)
}

# displaying means and bounds in dataframe
bootstrap.0 <- data.frame(means=means.group0, lower=lower.bounds0, upper=upper.bounds0)
bootstrap.0$idx <- 0
bootstrap.0
bootstrap.1 <- data.frame(means=means.group1, lower=lower.bounds1, upper=upper.bounds1)
bootstrap.1$idx <- 1
bootstrap.1
```

Produce a line chart that includes the bootstrapped mean and lower and upper percentile intervals for each group. Add appropriate labels and a legend. (6 points)

```{r, cache=TRUE, tidy=TRUE}
# combine two data frames
bootstrap <- rep(1:10,2) # boostrap number 1-10 for each treatment group
all.data <- cbind(bootstrap, rbind(bootstrap.0, bootstrap.1))
all.data

# line chart with bootstrap mean and lower/upper percentile intervals for each group
p <- ggplot(data=all.data, aes(x=bootstrap, y=means)) +
  scale_fill_manual(values=c("blue", "red"), name="Confidence Interval") +
  scale_colour_manual(values=c("blue", "red"), name="Mean") +
  geom_ribbon(aes(ymin=all.data$lower, ymax=all.data$upper, group=factor(idx),
                  fill=factor(idx)), alpha=0.4) +
  geom_line(aes(group=factor(idx), colour=factor(idx))) +
  ylab("Mean Outcome") + xlab("Bootstrap Number") + ggtitle("Bootstrapped Mean Outcome and CI by Group") + ylim(55,70)
p
```

# Question 4

Programming with classes. The following function will generate random patient information.

```{r, cache=TRUE, tidy=TRUE}
makePatient <- function() {
  vowel <- grep("[aeiou]", letters)
  cons <- grep("[^aeiou]", letters)
  name <- paste(sample(LETTERS[cons], 1), sample(letters[vowel], 1), sample(letters[cons], 1), sep='')
  gender <- factor(sample(0:1, 1), levels=0:1, labels=c('female','male'))
  dob <- as.Date(sample(7500, 1), origin="1970-01-01")
  n <- sample(6, 1)
  doa <- as.Date(sample(1500, n), origin="2010-01-01")
  pulse <- round(rnorm(n, 80, 10))
  temp <- round(rnorm(n, 98.4, 0.3), 2)
  fluid <- round(runif(n), 2)
  list(name, gender, dob, doa, pulse, temp, fluid)
}
```

1. Create an S3 class medicalRecord for objects that are a list with the named elements name, gender, date_of_birth, date_of_admission, pulse, temperature, fluid_intake. Note that an individual patient may have multiple measurements for some measurements. Set the RNG seed to 8 and create a medical record by taking the output of makePatient. Print the medical record, and print the class of the medical record. (5 points)

```{r, cache=TRUE, tidy=TRUE}
set.seed(8)

# creation of S3 class medicalRecord
new.patient <- makePatient()
names(new.patient) <- c("name", "gender", "dob", "doa", "pulse", "temp", "fluid")

class(new.patient)
class(new.patient) <- "medicalRecord" # set class to medicalRecord
new.patient
class(new.patient)
```

2. Write a medicalRecord method for the generic function mean, which returns averages for pulse, temperature and fluids. Also write a medicalRecord method for print, which employs some nice formatting, perhaps arranging measurements by date, and plot that generates a composite plot of measurements over time. Call each function for the medical record created in part 1. (5 points)

```{r, cache=TRUE, tidy=TRUE}
# medicalRecord method for generic function mean
# rounded 2 decimal places
mean.medicalRecord <- function(patient) {
  pulses <- as.numeric(unlist(patient["pulse"]))
  temps <- as.numeric(unlist(patient["temp"]))
  fluids <- as.numeric(unlist(patient["fluid"]))
  mean.pulses <- mean(pulses)
  mean.temps <- mean(temps)
  mean.fluids <- mean(fluids)
  cat(sprintf("Average Pulse: %.2f \nAverage temperature: %.2f \nFluid: %.2f", 
              mean.pulses, mean.temps, mean.fluids), "\n")
}

mean.medicalRecord(new.patient)

# medicalRecord method for print

print.patient <- function(patient) {
  name <- as.character(unlist(patient["name"]))
  gender <- as.character(unlist(patient["gender"]))
  date.birth <- as.Date(unlist(patient["dob"]), origin = "1970-01-01")
  dates.admitted <- as.Date(unlist(patient["doa"]), origin = "1970-01-01")
  pulses <- as.numeric(unlist(patient["pulse"]))
  temps <- as.numeric(unlist(patient["temp"]))
  fluids <- as.numeric(unlist(patient["fluid"]))
  patient.data <- data.frame(dates.admitted, pulses, temps, fluids)
  ordered.data <- patient.data[order(as.Date(patient.data$dates, "%d/%m/%Y"), decreasing = FALSE),] 
  cat(sprintf("Patient name: %s \nGender: %s \nDate of Birth: %s \n", name, gender, date.birth))
  print(ordered.data)
}

print.patient(new.patient)

# plotting of measurements
plot.patient <- function(patient) {
  dates.admitted <- as.Date(unlist(patient["doa"]), origin = "1970-01-01")
  pulses <- as.numeric(unlist(patient["pulse"]))
  temps <- as.numeric(unlist(patient["temp"]))
  fluids <- as.numeric(unlist(patient["fluid"]))
  patient.data <- data.frame(dates.admitted, pulses, temps, fluids)
  ordered.data <- patient.data[order(as.Date(patient.data$dates, "%d/%m/%Y"), decreasing = FALSE),] 
  par(mfrow=c(1,3))
  plot(x = ordered.data$dates.admitted, y = ordered.data$pulses, type = 'l', main = 
         "Patient Pulses Over Time", xlab = "Dates Admitted", ylab = "Pulses", col = "cornflower blue")
  plot(x = ordered.data$dates.admitted, y = ordered.data$temps, type = 'l', main = "Patient 
       Temperatures Over Time", xlab = "Dates Admitted", ylab = "Temperatures", col = "slate blue")
  plot(x = ordered.data$dates.admitted, y = ordered.data$fluids, type = 'l', main = "Patient 
       Fluid Intake Over Time", xlab = "Dates Admitted", ylab = "Fluid Intake", col = "medium sea green")
}

plot.patient(new.patient)

```

3. Create a further class for a cohort (group) of patients, and write methods for mean and print which, when applied to a cohort, apply mean or print to each patient contained in the cohort. Hint: think of this as a "container" for patients. Reset the RNG seed to 8 and create a cohort of ten patients, then show the output for mean and print. (5 points)

```{r, cache=TRUE, tidy=TRUE}
set.seed(8)

# class for cohort of 10 patients
cohort <- replicate(10, makePatient())
rownames(cohort) <- c("name", "gender", "dob", "doa", "pulse", "temp", "fluid")

# mean method for cohort
mean.10 <- function(patients) {
  mean.pulses <- vector()
  mean.temps <- vector()
  mean.fluids <- vector()
  for (i in 1:10) {
    mean.pulses[i] <- mean(as.numeric(unlist(patients[5,i])))
    mean.temps[i] <- mean(as.numeric(unlist(patients[6,i])))
    mean.fluids[i] <- mean(as.numeric(unlist(patients[7,i])))
  }
  means.all <- data.frame(mean.pulses, mean.temps, mean.fluids)
  print(means.all)
}

mean.10(cohort) # output of mean

print.10 <- function(patients) {
  for (i in 1:10) {
    name <- as.character(unlist(patients["name",i]))
    gender <- as.character(unlist(patients["gender",i]))
    date.birth <- as.Date(unlist(patients["dob",i]), origin = "1970-01-01")
    dates.admitted <- as.Date(unlist(patients["doa",i]), origin = "1970-01-01")
    pulses <- as.numeric(unlist(patients["pulse",i]))
    temps <- as.numeric(unlist(patients["temp",i]))
    fluids <- as.numeric(unlist(patients["fluid",i]))
    patient.data <- data.frame(dates.admitted, pulses, temps, fluids)
    ordered.data <- patient.data[order(as.Date(patient.data$dates, "%d/%m/%Y"), decreasing = FALSE),]
    cat(sprintf("\n"))
    cat(sprintf("Patient name: %s \nGender: %s \nDate of Birth: %s \n", name, gender, date.birth))
    print(ordered.data)
  }
}

print.10(cohort) # output of print

```
